
---
title: "Cover Type Classification with Random Forest and SVMLinear"
output:
  html_document:
    pandoc_args: [
      "+RTS", "-K64m",
      "-RTS"
    ]
---

<br><br>

__This code predicts forest categories using some cartographic variables.__

The dataset was taken from the UCI Machine learning repository. More details [here](https://archive.ics.uci.edu/ml/datasets/Covertype). 

This [link](https://archive.ics.uci.edu/ml/machine-learning-databases/covtype/covtype.data.gz) contains the full dataset. Download and unzip it in your project.

<br>
<h3> Function to clean data: </h3>


```{r message=FALSE, warning=FALSE}

library(caret)
library(randomForest)


cleanData <- function (d) {
  
  typecover=factor(d$Cover_Type)
  d <- cbind( d[,1:54], typecover)
  
  soilvars <- paste0('Soil_Type', 1:40)
  areavars <- paste0('Wilderness_Area', 1:4)
  nonattributes <- c(soilvars, areavars, 'Id', 'Cover_Type')
  
  # Convert soil variables into a single factor variable
  soils <- mapply(`*`, d[,soilvars], 1:40)
  soil <- factor(Reduce(`+`, data.frame(soils)), levels=1:40)
  
  # Convert wilderness area variables into a single factor variable
  areas <- mapply(`*`, d[,areavars], 1:4)
  area <- factor(Reduce(`+`, data.frame(areas)), levels=1:4)
  
  #subset(d, select=-c(Id, Cover_Type, soilvars, areavars))
  atts <- subset(d, select=setdiff(names(d), nonattributes))
  atts
  cbind(atts, soil=soil, area=area)
}
```

<br>
<h2> Random Forest </h2>
<h3> Read and split data into a train and test set: </h3>


```{r message=FALSE, warning=FALSE}

set.seed(2); 
raw_data <- read.csv('covtype.data')

trainingdata <- cleanData(raw_data)


inTrain <- createDataPartition(y = trainingdata[,11], p = .75,list = FALSE);

training <- trainingdata[ inTrain,];
testing <- trainingdata[-inTrain,];

labels_training <- training[,11]
training <- training[,-11]

labels_testing <- as.factor(testing[,11])
testing  <- testing[,-11]


#Takes only 25000 instances for the training
indices <- sample( 1:nrow( training ), 25000 )
training2 <- training[ indices, ]
labels_training2 <- labels_training[ indices ]

```

<br>
<h3> Training: </h3>


```{r warning=FALSE}
model <- randomForest(x=training2,y=labels_training2)
```

<br>
<h3> Evaluation: </h3>


```{r warning=FALSE}
classesPredictionModel <- predict(model, newdata = testing)
#classesPredictionModel
confusionMatrix(data = classesPredictionModel, labels_testing)
```

<br><br>
<h2> SVM </h2>
<h3> Split data into a train and test set: </h3>

```{r warning=FALSE}
svm_inTrain <- createDataPartition(y = raw_data[,55], p = .75,list = FALSE);
svm_training <- raw_data[ inTrain,];
svm_testing <- raw_data[-inTrain,];

svm_labels_training <- as.factor(svm_training[,55])
svm_training <- svm_training[,-55]

svm_labels_testing <- as.factor(svm_testing[,55])
svm_testing  <- svm_testing[,-55]

#Muestrea 25000 instancias del entrenamiento
svm_indices <- sample( 1:nrow( svm_training ), 25000 )
svm_training2 <- svm_training[ svm_indices, ]
svm_labels_training2 <- svm_labels_training[ svm_indices ]
```

<br>
<h3> Training: </h3>
```{r warning=FALSE,message=FALSE}
svmLinearmodel <- train(x=svm_training2[,1:54],y=svm_labels_training2, method = "svmLinear", tuneLength = 3);
```

<br>
<h3> Evaluation: </h3>
```{r warning=FALSE}
classesPredictionSVMLin <- predict(svmLinearmodel, newdata = svm_testing[,1:54])
confusionMatrix(data = classesPredictionSVMLin, svm_labels_testing)
```



